\documentclass[10pt,twocolumn,letterpaper]{article}

%% ── Packages ──────────────────────────────────────────────────────────────
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=0.75in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{tabularx}
\usepackage{float}
\usepackage{microtype}

%% ── Colors ────────────────────────────────────────────────────────────────
\definecolor{cascadeblue}{HTML}{0088FF}
\definecolor{cascadedark}{HTML}{001824}
\definecolor{codebg}{HTML}{F5F5F5}
\hypersetup{
  colorlinks=true,
  linkcolor=cascadeblue,
  urlcolor=cascadeblue,
  citecolor=cascadeblue
}

%% ── Listings ──────────────────────────────────────────────────────────────
\lstset{
  backgroundcolor=\color{codebg},
  basicstyle=\ttfamily\scriptsize,
  breaklines=true,
  frame=single,
  rulecolor=\color{gray!30},
  columns=fullflexible,
  keepspaces=true,
  xleftmargin=4pt,
  xrightmargin=4pt,
  aboveskip=6pt,
  belowskip=6pt
}

%% ── Section formatting ───────────────────────────────────────────────────
\titleformat{\section}{\large\bfseries}{\thesection.}{0.5em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{0.5em}{}
\titlespacing*{\section}{0pt}{10pt}{4pt}
\titlespacing*{\subsection}{0pt}{8pt}{3pt}
\setlength{\columnsep}{0.25in}

%% ── Header/Footer ────────────────────────────────────────────────────────
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\textsc{CASCADE}}
\fancyhead[R]{\small National Space Hackathon 2026}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0.4pt}

%% ── Title ─────────────────────────────────────────────────────────────────
\title{
  \vspace{-1.5em}
  {\LARGE\bfseries CASCADE: Autonomous Orbital Collision\\[2pt]
  Avoidance for Congested LEO}\\[6pt]
  {\normalsize National Space Hackathon 2026 --- IIT Delhi}
}
\author{
  Team Space\_Invaders\\[2pt]
  \small \url{https://github.com/Aqwerty321/PROJECT-B.O.N.K}
}
\date{March 2026}

\begin{document}
\maketitle
\thispagestyle{fancy}

%% ═══════════════════════════════════════════════════════════════════════════
\begin{abstract}
\noindent
Low Earth Orbit is entering an era of unprecedented congestion: over 27{,}000
tracked objects share an increasingly crowded shell, and every new
mega-constellation launch raises the probability of a cascading Kessler
event.  This report presents \textbf{CASCADE} (\textbf{C}ollision
\textbf{A}voidance \textbf{S}ystem for \textbf{C}onstellation
\textbf{A}utomation, \textbf{D}etection \& \textbf{E}vasion), a fully
autonomous, deterministic constellation manager built from the ground up to
address all six evaluation axes of the National Space Hackathon~2026 problem
statement---safety, fuel efficiency, constellation uptime, algorithmic
speed, visualization quality, and code integrity.  CASCADE ingests
high-frequency ECI~J2000 telemetry, screens conjunctions through a
mathematically conservative multi-tier filter cascade with a provable
zero-false-negative guarantee, and autonomously plans and schedules evasion
and recovery burns under realistic propulsion, cooldown, and
line-of-sight constraints across a six-station global ground network.
Under the reference scenario (50~satellites, 10{,}000~debris), the C++20
engine completes a full simulation tick---propagation, screening, CDM
generation, maneuver planning, and recovery scheduling---in
\textbf{13.3\,ms} mean wall-clock time.  Validated against a live
Space-Track OMM catalog of \textbf{27{,}191}~LEO objects, CASCADE maintains
zero missed conjunctions, sub-kilometer station-keeping accuracy, and
Tsiolkovsky-consistent fuel accounting across all test families.
The system ships as a single Docker image (\texttt{ubuntu:22.04},
port~8000) with a React~19 / TypeScript mission-console frontend built on
Canvas~2D and Three.js~WebGL, backed by 8~core safety gates integrated into
the Docker build and GitHub Actions workflow.
\end{abstract}

%% ═══════════════════════════════════════════════════════════════════════════
\section{Introduction}\label{sec:intro}

The accelerating growth of resident space objects in LEO---driven by
commercial mega-constellations, defunct payloads, and fragmentation
debris---has made autonomous collision avoidance a critical capability
for any responsible constellation operator.  The National Space
Hackathon~2026 distils this real-world challenge into a well-defined
engineering problem: build an Autonomous Constellation Manager (ACM) that
can (i)~ingest high-frequency ECI~J2000 state vectors, (ii)~predict
conjunctions up to 24~hours ahead without $O(N^{2})$ brute-force
enumeration, (iii)~autonomously compute, validate, and schedule evasion
and station-recovery maneuvers, and (iv)~present the full operational
picture through an interactive mission dashboard.

From the outset, our team adopted a design philosophy rooted in three
principles that guided every architectural decision:

\begin{itemize}[nosep,leftmargin=*]
  \item \textbf{Mathematical safety guarantees.}  A single missed
    conjunction can trigger a catastrophic debris cascade.  CASCADE is
    engineered so that every uncertain, ambiguous, or computationally
    degraded object pair is \emph{escalated} to a more expensive solver,
    never silently dropped.  The result is a provable zero-false-negative
    property across the entire filter cascade.

  \item \textbf{Deterministic, auditable execution.}  Identical inputs
    always produce the same maneuver sequence, the same CDM records, and
    the same fuel expenditure.  This determinism is not incidental---it is
    enforced by 8~standalone CTest gate binaries that run on every commit,
    making the safety guarantee reproducible and CI-verifiable.

  \item \textbf{Operational realism.}  The engine respects the physical
    constraints of real spacecraft operations: impulsive-burn propulsion
    with Tsiolkovsky fuel accounting, 600\,s thruster cooldown, 15\,m/s
    per-burn thrust limits, line-of-sight validation against a
    geographically distributed ground network, and a 10\,s signal-latency
    budget that forces proactive burn scheduling before blackout windows.
\end{itemize}

\noindent
The remainder of this report presents the system architecture
(\S\ref{sec:arch}), our engineering rationale for key design choices
(\S\ref{sec:decisions}), the numerical methods underpinning propagation
and screening (\S\ref{sec:methods}), the mission-console frontend
(\S\ref{sec:frontend}), validation evidence (\S\ref{sec:validation}),
and the deployment pipeline (\S\ref{sec:deploy}).

%% ═══════════════════════════════════════════════════════════════════════════
\section{System Architecture}\label{sec:arch}

CASCADE is a single-process C++20 engine with a deliberately minimal
dependency footprint: \texttt{cpp-httplib}~v0.15.3 for the REST server,
\texttt{simdjson}~v3.9.4 for streaming JSON ingestion, and Boost headers
for version verification---all managed through CMake FetchContent with no
system-level library requirements beyond the GCC~12 standard library and
\texttt{libgomp} for OpenMP.  This dependency discipline ensures that the
build is fully reproducible from a clean Ubuntu~22.04 base image.

Figure~\ref{fig:pipeline} illustrates the data path traversed by every
simulation tick, from telemetry ingestion through maneuver execution.

\begin{figure}[h]
\centering
\fbox{\parbox{0.95\linewidth}{\small\ttfamily
\begin{tabbing}
POST /api/telemetry \\
\quad $\downarrow$ \\
{[State Store]} -- lock-free SoA upsert \\
\quad $\downarrow$ \\
{[Propagation]} -- adaptive RK4+J2 (OpenMP) \\
\quad $\downarrow$ \\
{[Broad Phase]} -- $O(N \log N)$ SMA+incl bins \\
\quad $\downarrow$ \\
{[Narrow Phase]} -- MOID + TCA + RK4 refine \\
\quad $\downarrow$ \\
{[CDM Engine]} -- 24\,h predictive scan \\
\quad $\downarrow$ \\
{[COLA Planner]} -- RTN evasion + Tsiolkovsky \\
\quad $\downarrow$ \\
{[Recovery]} -- CW/ZEM slot-targeting burn \\
\quad $\downarrow$ \\
REST API + Frontend (port 8000)
\end{tabbing}
}}
\caption{End-to-end simulation tick pipeline.  Each stage is executed
  sequentially within a single tick; the engine guarantees that all
  propagation completes before screening begins, and all screening
  completes before maneuver decisions are made.}
\label{fig:pipeline}
\end{figure}

\subsection{Concurrency Model}

The engine employs a \emph{command-queue} concurrency pattern.
Mutating operations (telemetry ingestion, simulation steps, maneuver
scheduling) are serialized through a dedicated worker thread backed by
\texttt{std::future}/\texttt{std::promise} pairs, ensuring that no two
commands can interleave state modifications.  Read-only endpoints
(\texttt{GET /api/visualization/snapshot}, \texttt{GET /api/status})
operate concurrently via atomically published view pointers, avoiding
lock contention during high-frequency frontend polling.  Within a
simulation step, computationally intensive loops---orbital propagation and
narrow-phase refinement---are parallelized with OpenMP
(\texttt{\#pragma omp parallel for} with reduction clauses).

\subsection{State Representation}

All objects are stored in a Structure-of-Arrays (SoA) layout indexed by
string identifier, pre-allocated for 10{,}100~objects (50~satellites +
10{,}000~debris + margin).  Each record holds:

\begin{itemize}[nosep,leftmargin=*]
  \item 6-DOF ECI~J2000 state vector $(\vec{r}, \vec{v})$ in km and km/s,
    with per-object telemetry epoch for asynchronous ingestion
  \item Precomputed Keplerian elements $(a, e, i, \Omega, \omega, M)$
    derived via closed-form ECI-to-elements conversion at ingestion time
  \item Radial bounds $(r_{\min}, r_{\max})$ for shell overlap screening
  \item Fuel mass, cooldown timer, slot reference (bootstrapped from first
    telemetry and J2-secularly maintained), and full burn history
\end{itemize}

\subsection{Telemetry Ingestion}

Incoming telemetry arrives as a JSON payload containing an ISO-8601
timestamp and an array of objects, each with an identifier, type
(\texttt{SATELLITE} or \texttt{DEBRIS}), and 3-component position and
velocity vectors.  The \texttt{simdjson} on-demand parser streams the
payload through a pre-allocated padded buffer, avoiding intermediate
copies.  Invalid objects increment a diagnostic counter but never abort
the batch---operational resilience demands that a single malformed record
does not block ingestion of the remaining fleet.  Parsed objects are
applied atomically to the state store, and first-contact satellites
trigger automatic Keplerian element computation and slot-reference
bootstrapping.

\subsection{API Surface}

The engine exposes five primary endpoints on port~8000, bound to
\texttt{0.0.0.0} as required by the problem statement:

\begin{table}[h]
\centering\small
\begin{tabularx}{\linewidth}{@{}lX@{}}
\toprule
\textbf{Endpoint} & \textbf{Purpose} \\
\midrule
\texttt{POST /api/telemetry}       & Batch ECI state ingestion \\
\texttt{POST /api/maneuver/schedule} & Evasion/recovery burn submission \\
\texttt{POST /api/simulate/step}    & Advance simulation by $\Delta t$\,s \\
\texttt{GET /api/visualization/snapshot} & Geodetic snapshot for frontend \\
\texttt{GET /api/status}            & Engine health, diagnostics, filter stats \\
\bottomrule
\end{tabularx}
\caption{Primary REST API endpoints.}
\label{tab:api}
\end{table}

\noindent
A rich set of debug endpoints (\texttt{/api/debug/\{burns, conjunctions,
propagation, conflicts\}}) exposes internal ring buffers and per-tick
statistics, providing full operational observability without additional
logging infrastructure.

%% ═══════════════════════════════════════════════════════════════════════════
\section{Engineering Rationale}\label{sec:decisions}

Building a safety-critical orbital system under hackathon time constraints
requires disciplined decision-making at every stage.  Rather than
committing early to a single architecture, we deliberately structured the
development process around rapid prototyping, empirical evaluation, and
principled selection---investing time upfront in exploring alternatives so
that the final system rests on validated engineering judgments rather than
assumptions.  This section documents the six most consequential decisions
and the reasoning behind each.

\subsection{Prototyping with Julia, Deploying in Pure C++}

We recognized early that an orbital collision system involves two distinct
workloads: (i)~an \emph{offline tuning} problem, where multi-objective
optimizers sweep a large parameter space to find safe, efficient
configurations, and (ii)~a \emph{real-time inference} problem, where the
engine must execute a complete tick---propagation, screening, maneuver
planning---within a deterministic latency budget.  Julia's scientific
computing ecosystem (\texttt{DifferentialEquations.jl}, native BLAS,
first-class optimization libraries) made it the natural choice for
prototyping the tuner, so we built a complete NSGA-II parameter optimizer
via the \texttt{jluna} C++/Julia bridge, driving 69~engine parameters
across safety, fuel, and latency objectives.

Once the tuner proved that the parameter space could be safely explored, we
made a deliberate deployment decision: the production runtime would be pure
C++20 with zero Julia dependency.  The reasoning was threefold.  First, the
problem statement specifies a cold \texttt{docker build} on the grading
VM, and the Julia~1.10 runtime adds $\sim$200\,MB to the image plus 4--8\,s
of JIT warm-up---unacceptable for a latency-sensitive submission.  Second,
every numerical kernel the engine needs (RK4 integration, Keplerian solvers,
CW state-transition matrices) admits a compact, dependency-free C++
implementation that compiles to tight machine code under \texttt{-O2}.
Third, separating tuning from inference allowed us to guard the Julia
bridge behind a CMake option (\texttt{PROJECTBONK\_ENABLE\_JULIA\_RUNTIME})
and ship a minimal binary that links only against \texttt{libgomp1}---the
OpenMP runtime---and the C++ standard library.

The offline tuner remains a first-class tool: its NSGA-II runs (20
population $\times$ 5~generations, 120~evaluations) produced the
production defaults that the engine ships with, and all~69 parameters
remain overridable at runtime via \texttt{PROJECTBONK\_*} environment
variables, enabling parameter sweeps without recompilation.

\subsection{Staged Filter Activation via Shadow Deployment}

Aerospace-grade conjunction screening systems face an inherent tension:
aggressive filtering reduces computational cost, but every new filter
introduces a risk of inadvertently discarding a true conjunction.  In
mission-critical contexts, releasing a new filter \emph{directly into
production} is analogous to deploying untested flight software---the
consequences of a false negative are catastrophic.

We resolved this tension by adopting a \emph{shadow deployment} discipline
inspired by industry practices in fault-tolerant systems.  Every new
screening filter is introduced in three stages:

\begin{enumerate}[nosep,leftmargin=*]
  \item \textbf{Shadow mode}: The filter evaluates every candidate pair
    and logs its rejection decision, but the pair is \emph{not} discarded.
    The full downstream pipeline processes the pair regardless.
  \item \textbf{Calibration}: Shadow-mode telemetry---exposed via
    \texttt{/api/status?details=1}---accumulates rejection statistics
    against the ground-truth RK4 reference.  The team reviews false-negative
    rates, threshold sensitivity, and edge-case behavior.
  \item \textbf{Activation}: Only after the filter demonstrates zero false
    negatives across diverse orbital regimes (circular, eccentric,
    sun-synchronous, crossing) is it promoted from shadow to hard-reject
    mode via environment variable.
\end{enumerate}

This pattern is currently applied to the D-criterion broad-phase gate,
the plane/phase angular gate, and the MOID proxy.  It guarantees that the
zero-false-negative invariant is maintained at every point in the
development timeline, while still allowing the team to develop and
validate increasingly aggressive filters that will reduce narrow-phase
compute by an estimated 40--60\% upon activation.

\subsection{Clohessy--Wiltshire Recovery with Heuristic Safeguard}

Station recovery after an evasion maneuver is a classical relative-motion
problem.  We evaluated two approaches from the start: a proportional RTN
gain controller (simple, robust, but producing suboptimal $\Delta V$
vectors on eccentric orbits) and a full Clohessy--Wiltshire (CW)
state-transition inversion (mathematically optimal for near-circular
relative motion, but singular at certain CW horizons).

The CW formulation computes the recovery impulse by inverting the
in-plane $2 \times 2$ velocity-to-position submatrix of the full
$6 \times 6$ state-transition matrix.  We blend this
with velocity-error damping:
%
\begin{equation}
  \Delta\vec{v}_{\text{rec}} = \alpha\,\Phi_{rv}^{-1}\,\delta\vec{r}
    + \beta\,\delta\vec{v}
\end{equation}
%
with $\alpha = 0.70$ (position correction weight) and
$\beta = 0.30$ (velocity damping), tuned by the NSGA-II optimizer.  The
CW horizon is set to $0.25 \times T_{\text{orbit}}$, clamped to
$[300, 5400]$\,s.  We further cap the CW correction at 15\% of remaining
$\Delta V$ budget and 40\% of the heuristic norm to prevent
over-correction---a nuance that matters when recovery burns occur near
orbital resonances where the CW matrix condition number degrades.

The proportional controller is retained as an automatic fallback for the
specific regime where the CW matrix becomes near-singular ($nt \approx
k \cdot 2\pi$).  This dual-solver architecture ensures that recovery
planning never stalls, while defaulting to the mathematically superior
solution in the vast majority of cases.  The regression gate
\texttt{recovery\_slot\_gate} enforces
$\Delta_{\text{slot}} \leq 0.08$~deterministically across all test
configurations.

\subsection{Operational Dashboard over Visual Spectacle}

The frontend design was driven by a careful reading of the problem
statement's evaluation criteria.  Section~7 allocates 15\% weight to
\emph{UI/UX \& Visualization} and specifically requires a ground track
map, conjunction bullseye plot, telemetry heatmaps, and maneuver
timeline---all inherently 2D operational surfaces.  We recognized that
the value proposition of the frontend lies in \emph{situational
awareness}: enabling a mission operator to understand fleet status,
conjunction threats, and burn schedules at a glance.

Accordingly, we built the primary dashboard around Canvas~2D Mercator
ground tracks with 90-minute historical trails and forecast paths, a
polar ``bullseye'' threat-assessment view with radial time-to-closest-%
approach encoding, a scrollable Gantt maneuver timeline with blackout
overlays, and per-satellite fuel gauges.  A Three.js WebGL globe provides
supplementary 3D spatial context, but it supports rather than dominates
the interface.  This architecture ensures that every pixel on screen serves
an operational purpose aligned with the judge rubric, delivering the kind
of mission-console experience that a real spacecraft operator would
recognize.  The current shell reinforces that principle by keeping shared
controls compact: simulation stepping lives in the top command strip,
while global focus and the operator checklist move into a collapsible
sidebar so route-level canvases remain visually dominant.

\subsection{Robust MOID Screening via Multi-Mode Solver}

Minimum Orbit Intersection Distance (MOID) is a powerful pre-filter:
if two orbits never come within a threshold distance of each other
\emph{anywhere} along their trajectories, no time-domain refinement is
necessary.  The challenge is that analytical MOID solvers---such as the
Gronchi~(2005) quartic formulation we implemented in
\texttt{orbit\_math.cpp}---are sensitive to orbital element degeneracies.
Near-circular orbits (where the argument of perigee is undefined) and
near-equatorial orbits (where RAAN is poorly determined) produce
ill-conditioned systems.

Rather than restricting the analytical solver to well-conditioned regimes
and leaving degenerate cases un-screened, we designed a \emph{multi-mode}
MOID architecture:

\begin{itemize}[nosep,leftmargin=*]
  \item \textbf{PROXY}: A $72 \times 72$ true-anomaly grid (5{,}184
    evaluations per pair) that is robust to all orbit types, including
    circular and equatorial.
  \item \textbf{HF (High-Fidelity)}: Coarse global sampling at
    $2\times$ resolution followed by 3~iterations of local refinement
    ($5 \times 5$ grid, halving step size), with pre-computed satellite
    orbit positions to amortize trigonometric evaluations.
  \item \textbf{ANALYTICAL}: The full ellipse-to-ellipse minimum
    distance computation for well-conditioned pairs.
\end{itemize}

The grid resolution (72~samples per orbit) was calibrated through
analysis: at LEO altitudes, $5^{\circ}$~angular resolution corresponds
to $\sim$50\,km arc length, providing reliable sub-kilometer MOID
estimates.  The 2.0\,km reject threshold ensures that only genuinely
non-intersecting orbit pairs are discarded.  The PROXY mode is deployed
in shadow mode, consistent with our staged activation discipline.

\subsection{Provably Safe Fail-Open Escalation}

When propagation fails for an object---producing NaN states, divergent
trajectories, or eccentricities that exceed the valid domain---the engine
must decide how to screen that object's conjunctions.  An intuitive
approach would be a tiered fallback: use last-known elements for a
reduced broad-phase check.  We formally analyzed this option and
identified a critical flaw: if an object's orbital elements are stale or
invalid, the broad phase cannot guarantee that its shell-overlap test
captures all crossing-orbit debris.  A stale semi-major axis could shift
an object into the wrong SMA bin, causing it to miss pairs that a correct
binning would catch.

The only strategy that \emph{provably} maintains zero false negatives is
to treat propagation-failed objects as maximally uncertain: the engine
bypasses all cheap filters and checks the failed object against the
\emph{entire} debris catalog using the full narrow-phase pipeline.  This
is computationally expensive---$O(M)$ narrow-phase evaluations for each
failed object---but the cost is bounded (propagation failures are rare in
practice, typically affecting $<$0.1\% of objects per tick) and the
safety guarantee is absolute.  Detailed fail-open counters are exposed
in the engine's status API, providing full transparency into when and why
this escalation path activates.

%% ═══════════════════════════════════════════════════════════════════════════
\section{Numerical Methods}\label{sec:methods}

\subsection{Orbital Propagation}

Propagating 10{,}000+ objects per tick demands a scheme that is
simultaneously accurate enough to detect 100\,m-scale conjunctions and
fast enough to fit within a deterministic latency budget.  We designed an
\emph{adaptive three-mode} propagator that allocates computational effort
in proportion to each object's orbital complexity.

\paragraph{Mode 1 --- J2 Secular Fast Path.}
The vast majority of LEO objects occupy near-circular ($e \leq 0.02$),
high-perigee ($\geq 500$\,km) orbits where the dominant perturbation is
Earth's oblateness.  For these objects, we apply closed-form J2 secular
rates to mean elements:
%
\begin{align}
  \dot{\Omega} &= -\frac{3}{2} n J_2 \!\left(\frac{R_E}{p}\right)^{\!2} \cos i \\
  \dot{\omega} &= \frac{3}{4} n J_2 \!\left(\frac{R_E}{p}\right)^{\!2} (5\cos^{2}\!i - 1)
\end{align}
%
where $n = \sqrt{\mu / a^3}$, $p = a(1-e^2)$, and $R_E = 6378.137$\,km.
An extended fast lane admits $e \leq 0.003$, perigee $\geq 650$\,km,
$\Delta t \leq 45$\,s, covering the operating regime of most commercial
constellations.  Long horizons are sub-stepped at 300\,s intervals to
bound secular-rate linearity error.

\paragraph{Mode 2 --- Probe-Then-Escalate.}
Objects that fall outside the fast-lane criteria first receive a
J2-secular prediction, followed by a coarse RK4 probe (max step 120\,s)
that validates the result.  If the disagreement exceeds 0.5\,km in
position or 0.5\,m/s in velocity, the engine automatically escalates to
full numerical integration.  This two-stage check avoids paying the full
RK4 cost for objects that are well-approximated by secular theory, while
catching those that genuinely require numerical treatment.

\paragraph{Mode 3 --- Full RK4+J2.}
Objects with low perigee ($< 350$\,km), high eccentricity ($e \geq 0.98$),
or long propagation horizons ($\Delta t > 21{,}600$\,s) are integrated
with classical fourth-order Runge--Kutta, maximum substep 60\,s,
under the J2-perturbed acceleration:
%
\begin{equation}
  \vec{a}_{J2} = \frac{3}{2}\frac{J_2\,\mu\,R_E^2}{r^5}
  \begin{bmatrix}
    x(5z^2/r^2 - 1) \\
    y(5z^2/r^2 - 1) \\
    z(5z^2/r^2 - 3)
  \end{bmatrix}
\label{eq:j2}
\end{equation}
%
Finite-value guards after every substep prevent NaN propagation;
any divergent state triggers the fail-open escalation path described in
\S\ref{sec:decisions}.

\paragraph{Accuracy.}  Maximum position error at 86{,}400\,s
(a full orbit day) is \textbf{$<$\,0.061\,km}, verified by the
regression gate \texttt{phase2\_regression\_gate}---approximately
three orders of magnitude below the 100\,m collision threshold.

\paragraph{Why RK4 over higher-order adaptive integrators.}
We deliberately chose classical RK4 over RKF4(5) or the Dormand--Prince
pair.  The collision threshold (100\,m) is orders of magnitude larger than
RK4 truncation error at a 60\,s substep, so the additional accuracy of a
higher-order method provides no safety benefit.  More importantly, RK4
avoids the overhead of embedded error estimation and step-rejection logic,
keeping the per-object cost \emph{predictable}---a property that directly
enables deterministic tick budgets.  Our probe-then-escalate scheme
already provides architectural-level adaptivity where it
matters: allocating cheap secular propagation to easy objects and full
numerical integration to hard ones.

\subsection{Coordinate Transforms}

All internal computation uses ECI~J2000 coordinates.  For visualization
and ground-station LOS validation, the engine implements a full
coordinate-transform chain:

\begin{itemize}[nosep,leftmargin=*]
  \item \textbf{ECI $\to$ ECEF}: Rotation about the $z$-axis by the
    Greenwich Mean Sidereal Time angle, computed from the standard
    polynomial expansion (IAU~2000 compatible).
  \item \textbf{ECEF $\to$ Geodetic}: Iterative Bowring algorithm on the
    WGS-84 ellipsoid ($a = 6378.137$\,km, $f = 1/298.257223563$), converging
    to $10^{-13}$\,rad in at most 8~iterations.
  \item \textbf{Geodetic $\to$ ECEF}: Prime vertical radius of curvature
    $N(\phi)$ with WGS-84 constants for ground station positions.
  \item \textbf{Local ENU}: East-North-Up frame at each ground station for
    elevation-angle computation (see \S\ref{sec:los}).
\end{itemize}

\subsection{Broad-Phase Candidate Generation}

The broad phase reduces the $O(N^2)$ all-pairs space through three
sequential geometric filters:

\begin{enumerate}[nosep,leftmargin=*]
  \item \textbf{SMA--Inclination Band Indexing.}
    Objects are binned by semi-major axis ($\Delta a = 500$\,km) and
    inclination ($\Delta i = 20^{\circ}$).  Flat sorted arrays with binary
    search replace hash maps, exploiting cache locality on modern CPUs.
    Neighbor search extends $\pm 2$~bins in each dimension to handle
    boundary-crossing orbits.

  \item \textbf{Shell Overlap Test.}
    Perigee $r_{\min} = a(1-e)$ and apogee $r_{\max} = a(1+e)$ are
    compared with a 50\,km margin.  Objects with invalid elements fall back
    to an instantaneous radius $\pm 200$\,km envelope---an intentional
    over-approximation that preserves the zero-false-negative guarantee at
    the cost of passing more candidates to the narrow phase.

  \item \textbf{D-Criterion Gate.}
    A simplified Drummond proximity metric:
    %
    \begin{equation}
      D = \sqrt{(\Delta a / \bar{a})^2 + \Delta e^2
          + (2\sin(\Delta i / 2))^2}
    \end{equation}
    %
    with reject threshold $D > 2.0$.  Deployed in shadow mode as part of
    the staged activation discipline (\S\ref{sec:decisions}).
\end{enumerate}

\noindent
\textbf{Fail-open path.}  Objects with invalid elements,
$e > 0.2$, or radial span exceeding bin coverage bypass all cheap filters
and proceed directly to the narrow phase.  Detailed fail-open counters
are exposed in \texttt{/api/status?details=1}, providing transparency
into the fraction of the catalog that exercises this conservative path.

\subsection{Narrow-Phase Collision Screening}

Candidate pairs from the broad phase enter a multi-gate narrow pipeline
designed to progressively concentrate computational effort on the
highest-risk pairs:

\begin{enumerate}[nosep,leftmargin=*]
  \item \textbf{Plane/Phase Angular Gate.}
    Angular momentum vectors are computed from $(i, \Omega)$ and used to
    assess orbital plane alignment.  Pairs with orbital plane angle
    $> 75^{\circ}$ or mean anomaly phase separation $> 150^{\circ}$ are
    rejected---their orbits cannot geometrically intersect within the
    screening horizon.  Objects with $e > 0.2$ bypass this gate
    (fail-open), since their large radial excursion can bridge wide
    angular separations.

  \item \textbf{MOID Proxy.}
    The multi-mode MOID solver (\S\ref{sec:decisions}) evaluates minimum
    orbit-to-orbit distance.  In the default PROXY mode, 5{,}184 sampled
    evaluations per pair provide sub-kilometer accuracy across all orbital
    regimes.  Reject threshold: 2.0\,km.

  \item \textbf{Linear TCA Approximation.}
    Four simultaneous linear-segment minimum-distance estimates
    (start-velocity, end-velocity backward, averaged, and secant
    approximations) are computed over the step interval.
    The minimum $d^2$ across all four determines passage.

  \item \textbf{RK4 Micro-Refinement.}
    Pairs within 0.10\,km above the screening threshold receive a
    fine-grained RK4 propagation ($\Delta t_{\max} = 5$\,s), resolving
    any ambiguity introduced by the linear TCA approximation.

  \item \textbf{Full-Window RK4 Refinement.}
    Pairs within 0.20\,km receive a budget-limited, OpenMP-parallelized
    full-window refinement (16~trajectory samples, 1\,s substep).
    Thread-safe atomic budget decrement prevents compute runaway:
    budget tiers range from 20 (for $>$2M pairs) to 96 ($<$100K pairs),
    dynamically scaled by pair count and step size, clamped to $[8, 192]$.
\end{enumerate}

\noindent
\textbf{Uncertainty promotion.}  Hypervelocity encounters (relative speed
$>$\,8\,km/s) receive an additional 0.10\,km screening margin, reflecting
the reality that positional uncertainty grows faster at high closing
speeds.  The effective screening distance is:
%
\begin{equation}
  d_{\text{screen}} = \underbrace{0.100}_{\text{collision}} +
                      \underbrace{0.020}_{\text{TCA guard}} = 0.120 \;\text{km}
\end{equation}

\subsection{24-Hour Predictive CDM Generation}

After each simulation tick, the engine performs a forward-looking
conjunction scan over a configurable horizon (default 86{,}400\,s = 24\,h).
Broad-phase candidate pairs are independently RK4+J2 propagated from their
post-step states in 600\,s substeps (RK4 max internal step 30\,s).  For each
time sample, the engine evaluates miss distance and approach speed, generating
a \emph{Conjunction Data Message} (CDM) record that includes satellite/debris
identifiers, TCA epoch, miss distance, approach speed, both ECI positions,
collision flag, and severity classification.  Up to 1{,}024 predictive
records are retained per scan, sorted by severity (CRITICAL first, then
ascending miss distance), and exposed via the status and debug APIs.

Pairs where RK4 propagation fails---producing NaN or divergent states---are
conservatively classified as CRITICAL, triggering the autonomous COLA
pipeline.  This ensures that computational failures never suppress
collision warnings.

\subsection{Autonomous Maneuver Planning}

\subsubsection{COLA Planner}

When a predictive CDM record reaches CRITICAL severity ($d < 100$\,m),
the engine triggers an autonomous evasion burn search over a discrete
candidate grid:

\begin{itemize}[nosep,leftmargin=*]
  \item 4~burn-time fractions of the remaining lead time
  \item 4~$\Delta V$ scales: $\{0.5, 1, 2, 4\} \times 0.001$\,km/s
  \item 10~RTN directions: $\pm T$, $\pm R$, 4~diagonal $T/R$, $\pm N$
\end{itemize}

Each candidate burn is fully forward-propagated via RK4 with the impulse
applied, and the resulting trajectory is screened against all active debris
to compute the post-maneuver miss distance.  The RTN frame is
constructed using the true tangential direction:
\begin{equation}
  \hat{r} = \frac{\vec{r}}{|\vec{r}|}, \quad
  \hat{n} = \frac{\vec{r} \times \vec{v}}{|\vec{r} \times \vec{v}|}, \quad
  \hat{t} = \hat{n} \times \hat{r}
\end{equation}

\noindent
Candidates are ranked by a lexicographic multi-criteria comparator: safety
class (miss distance relative to 0.15\,km target) $\to$ $|\Delta V|$ $\to$
normal component magnitude $\to$ slot error $\to$ predicted miss distance.
A maximum of 3~CRITICAL CDMs are evaluated per satellite per tick,
bounding the worst-case planning cost.

\subsubsection{Fuel Accounting}

Propellant expenditure follows the Tsiolkovsky rocket equation:
%
\begin{equation}
  \Delta m = m_{\text{current}} \!\left(1 - e^{-|\Delta\vec{v}| / v_e}\right),
  \quad v_e = I_{\text{sp}} \cdot g_0
\label{eq:tsiolkovsky}
\end{equation}
%
with $I_{\text{sp}} = 300$\,s, $g_0 = 9.80665$\,m/s$^2$,
$m_{\text{dry}} = 500$\,kg, $m_{\text{fuel,init}} = 50$\,kg.
A 5\%~fuel guard (2.5\,kg) triggers automatic graveyard orbit transfer,
ensuring that end-of-life satellites are disposed of responsibly rather
than becoming debris themselves.

\subsubsection{Constraint Enforcement}

The engine enforces the full set of operational constraints specified in
the problem statement:

\begin{itemize}[nosep,leftmargin=*]
  \item \textbf{Cooldown}: 600\,s mandatory rest between burns on the same
    satellite, preventing thruster thermal overload.
  \item \textbf{Thrust cap}: $|\Delta V| \leq 15$\,m/s per burn,
    reflecting realistic chemical thruster limits.
  \item \textbf{LOS validation}: Upload windows are scanned in 20\,s steps
    across a 1{,}800\,s horizon against a 6-station global ground
    network (\S\ref{sec:los}).  A 10\,s signal latency is enforced.
  \item \textbf{Blackout handling}: Burns must be computed and queued
    before the satellite exits the last ground-station footprint,
    ensuring commands are uploaded in time.
\end{itemize}

\subsection{Ground Network and Line-of-Sight}\label{sec:los}

CASCADE models a globally distributed 6-station ground network, designed
to maximize coverage of polar, equatorial, and mid-latitude LEO orbits:

\begin{table}[h]
\centering\small
\begin{tabular}{@{}llrc@{}}
\toprule
\textbf{Station} & \textbf{Location} & \textbf{Min El.} \\
\midrule
ISTRAC Bengaluru     & $13.0^{\circ}$\,N, $77.5^{\circ}$\,E  & $5^{\circ}$  \\
Svalbard Satellite   & $78.2^{\circ}$\,N, $15.4^{\circ}$\,E  & $5^{\circ}$  \\
Goldstone Tracking   & $35.4^{\circ}$\,N, $116.9^{\circ}$\,W & $10^{\circ}$ \\
Punta Arenas         & $53.2^{\circ}$\,S, $70.9^{\circ}$\,W  & $5^{\circ}$  \\
IIT Delhi Ground Node& $28.5^{\circ}$\,N, $77.2^{\circ}$\,E  & $15^{\circ}$ \\
McMurdo Station      & $77.8^{\circ}$\,S, $166.7^{\circ}$\,E & $5^{\circ}$  \\
\bottomrule
\end{tabular}
\caption{Ground station network.  Elevation constraints are per-station to
  model real-world terrain and infrastructure masks.}
\label{tab:gs}
\end{table}

\noindent
For each candidate burn, the engine propagates the satellite's position to
the planned upload epoch augmenting it with J2 secular rates, converts from
ECI to ECEF via GMST rotation, transforms the satellite-to-station vector
into the local ENU frame, and computes the elevation angle:
%
\begin{equation}
  \theta_{\text{el}} = \arctan\!\left(\frac{u_{\text{Up}}}
    {\sqrt{u_{\text{East}}^2 + u_{\text{North}}^2}}\right)
\end{equation}
%
If $\theta_{\text{el}}$ exceeds the station's minimum elevation mask at
any point within the upload window, LOS is confirmed and the burn is
eligible for scheduling.  Station configurations are loaded from CSV at
runtime (\texttt{PROJECTBONK\_GROUND\_STATIONS\_CSV}), with a compiled-in
fallback array ensuring graceful degradation if the file is unavailable.

\subsection{Recovery Planning}

Every evasion maneuver triggers a paired recovery burn computed via the
CW/ZEM solver described in \S\ref{sec:decisions}.  The station-keeping
box is defined as a 10\,km spherical radius around the nominal orbital
slot, which is bootstrapped from first telemetry and propagated with
J2 secular rates.  Recovery burns are conservatively budgeted: only
5\% of the remaining $\Delta V$ allocation is consumed per recovery
request, preventing over-correction cascades.  Each planned recovery
trajectory is re-simulated through the conjunction screening pipeline---a
\emph{ghost recovery} simulation---to ensure that the recovery burn
itself does not create a new collision risk.

%% ═══════════════════════════════════════════════════════════════════════════
\section{Mission Console Frontend}\label{sec:frontend}

The frontend is a React~19 / TypeScript single-page application built with
Vite~6 and rendered via Canvas~2D and Three.js~WebGL.  During live
operation it polls the snapshot stream at 1\,Hz and the status, burn,
conjunction, and trajectory streams at 2\,s cadence.  The UI is organized
around a shared tactical shell plus six purpose-built route views:

\begin{enumerate}[nosep,leftmargin=*]
  \item \textbf{Command / Mission Theatre} (\texttt{\#/command}):
    Fleet-wide posture summary, shared mission-time context, a Three.js
    orbital globe for command geometry, and a right-hand mission rail for
    focus, threat, burn, and slot-integrity state.  This is the default
    landing view and the operator's main situational-awareness surface.

  \item \textbf{Ground Track} (\texttt{\#/track}):
    Mercator-projection canvas rendering of all satellite positions with
    90-minute historical trails and 90-minute forecast arcs, a live
    solar terminator overlay, ground-station markers with coverage
    footprints, and a persistent map legend with epoch timestamp.
    Geodetic coordinates are derived from the engine's ECI-to-geodetic
    transform chain.

  \item \textbf{Threat Assessment} (\texttt{\#/threat}):
    Polar ``bullseye'' plot centered on the selected satellite.  Radial
    distance encodes time-to-closest-approach; angular position encodes
    the approach vector in the satellite's local frame.  CDM severity
    determines marker color (green/yellow/red), while the encounter queue
    and detail rail expose the most urgent watch-lane events.

  \item \textbf{Burn Operations} (\texttt{\#/burn-ops}):
    Scrollable Gantt-style maneuver timeline with evasion and recovery
    burns, blackout-zone overlays, and cooldown friction markers.
    Per-satellite fuel gauges show remaining propellant and $\Delta V$
    budget.

  \item \textbf{Evasion Analysis} (\texttt{\#/evasion}):
    Counterfactual burn comparison, clearance-gain versus fuel-spent
    analysis, and operator-facing explanation of the avoidance trade space.

  \item \textbf{Fleet Status} (\texttt{\#/fleet-status}):
    Constellation-level analytics---fuel heatmaps across the fleet,
    $\Delta V$ expenditure versus collisions avoided, station-keeping
    drift metrics, and satellite availability tracking.
\end{enumerate}

Across these pages, the shell preserves a single cross-route focus model:
pinning a spacecraft in one view updates the command globe, threat queue,
track detail rail, and burn context everywhere else.  This shared-state
approach reduces duplicated controls while keeping route-level evidence
aligned to the same selected vehicle.

\noindent
The shared shell provides a persistent summary rail, a sticky truth and
freshness banner, inline simulation-step controls, reasoning / spotlight /
audio toggles, and a collapsible sidebar whose expanded utility panels house
the global focus console and operator checklist.  Snapshot, conjunction,
burn, and trajectory endpoints jointly supply the route-level context
consumed by these views.  Canvas~2D drives the ground-track, bullseye, and
Gantt surfaces; Three.js powers the 3D orbital theatre.  The entire
frontend is statically built, bundled into the Docker image, and served by
\texttt{cpp-httplib}'s mount-point mechanism---no separate web server
required.

%% ═══════════════════════════════════════════════════════════════════════════
\section{Validation and Evidence}\label{sec:validation}

A safety-critical system requires validation that is as rigorous as the
software it certifies.  CASCADE's test infrastructure is designed around a
principle borrowed from aerospace V\&V practice: every safety invariant is
backed by a standalone, deterministic gate that can be executed
independently and whose pass/fail verdict is unambiguous.

\subsection{Test Infrastructure}

When configured with \texttt{BUILD\_TESTING=ON}, the source tree registers
8~core CTest safety gates and builds the corresponding standalone binaries
in \texttt{tools/}.  Complementary shell harnesses in \texttt{scripts/}
exercise the same invariants in packaged-runtime scenarios.  The Dockerfile
runs \texttt{ctest} inside the builder stage---if any registered gate
fails, the image build itself fails---and GitHub Actions additionally runs
the dedicated gate scripts plus the Docker smoke test on pushes and pull
requests targeting \texttt{main}.

\begin{table}[h]
\centering\small
\begin{tabularx}{\linewidth}{@{}lX@{}}
\toprule
\textbf{Gate} & \textbf{Enforced Invariant} \\
\midrule
\texttt{phase2\_regression}        & Propagator drift $< 0.061$\,km at 86{,}400\,s \\
\texttt{phase2\_regression\_short} & Strict-adaptive stability at 1\,s / 120\,s horizon \\
\texttt{broad\_phase\_validation}  & Shell filter $\supseteq$ brute-force baseline \\
\texttt{narrow\_phase\_false\_neg.} & Zero missed collisions vs.\ RK4 reference \\
\texttt{recovery\_slot\_gate}      & Deterministic slot recovery \\
\texttt{recovery\_planner\_inv.}   & CW/ZEM convergence, slot error $\leq 0.08$ \\
\texttt{maneuver\_ops\_invariants} & Cooldown, fuel, thrust-cap enforcement \\
\texttt{api\_contract\_gate}       & PS-required API surface conforms to schema \\
\bottomrule
\end{tabularx}
\caption{Core safety gates exercised through CTest / CI.  Each gate is a
  binary verdict---pass or fail---with no subjective interpretation.}
\label{tab:gates}
\end{table}

\subsection{Real-Data Validation at Scale}

Synthetic test suites verify correctness under controlled conditions, but
real-world orbital catalogs exercise the engine against the full diversity
of orbit types, element qualities, and edge cases that exist in the actual
LEO population.  To validate CASCADE under realistic conditions, we
ingested a live Space-Track OMM catalog (\texttt{data.txt}, 34\,MB)
through the scenario generator (\texttt{tools/real\_data\_scenario\_gen}).

The generator parsed 30{,}000+ objects via simdjson, filtered to LEO
(periapsis $< 2{,}000$\,km altitude), classified payloads as satellites
and the remainder as debris, and ran simulation ticks at full scale.
Results:

\begin{itemize}[nosep,leftmargin=*]
  \item \textbf{27{,}191} LEO objects loaded in $<$\,1\,s with zero ECI
    conversion failures
  \item 10~conjunctions detected against the real catalog at 2.1\,s/tick
    (full 27K scale)
  \item All broad-phase filters correctly preserved pairs that the
    narrow phase subsequently confirmed
  \item Zero false negatives across the full catalog run
\end{itemize}

\noindent
This real-data exercise also served as a stress test for the fail-open
path: objects with poorly-determined elements in the Space-Track catalog
were automatically escalated to the full narrow-phase pipeline, exercising
the conservative screening path under production-representative conditions.

\subsection{Safety Score Evidence}

The problem statement weights safety at 25\%, the highest of all six
criteria, and imposes a severe penalty for even a single collision.
CASCADE's safety evidence rests on three pillars:

\begin{itemize}[nosep,leftmargin=*]
  \item \textbf{Zero false negatives} across 6~independent test families:
    broad-phase sanity, narrow-phase canary, false-negative gate,
    calibration probe, regression, and maneuver-ops.
  \item \textbf{Transparent fail-open accounting}: Every object with
    uncertain or invalid elements is escalated to the full narrow-phase
    pipeline.  Fail-open counters are tracked per tick and exposed in the
    status API, providing auditable evidence that no object is silently
    dropped.
  \item \textbf{Conservative screening margin}: Effective collision
    envelope of $0.120$\,km, with uncertainty promotion to $0.220$\,km for
    hypervelocity encounters ($>$\,8\,km/s relative speed).
\end{itemize}

\subsection{Fuel Efficiency Evidence}

\begin{itemize}[nosep,leftmargin=*]
  \item Tsiolkovsky-consistent fuel deduction validated against
    analytical expectations in \texttt{maneuver\_ops\_invariants\_gate}.
  \item CW/ZEM recovery solver produces lower $|\Delta V|$ than the
    proportional heuristic across all tested orbit configurations.
  \item Evasion fuel and recovery fuel are tracked independently in burn
    summaries, enabling per-category $\Delta V$ accounting.
  \item Recovery budgets are capped at 5\% of remaining $\Delta V$
    per request, ensuring that no single over-correction event depletes
    the constellation's margin.
\end{itemize}

\subsection{Algorithmic Speed Evidence}

\begin{table}[h]
\centering\small
\begin{tabular}{@{}lrl@{}}
\toprule
\textbf{Scenario} & \textbf{Mean Tick} & \textbf{Scale} \\
\midrule
50 sats / 10{,}000 debris   & 13.3\,ms & Baseline \\
100 sats / 20{,}000 debris  & 72.6\,ms & 2$\times$ stress \\
27{,}191 LEO objects (real)  & 2.1\,s   & Full Space-Track \\
\bottomrule
\end{tabular}
\caption{Tick latency benchmarks (\texttt{phase3\_tick\_benchmark}).}
\label{tab:perf}
\end{table}

\noindent
The broad phase reduces the candidate count from $O(N \cdot M)$ to a
small fraction via SMA/inclination binning.  OpenMP parallelization of
the narrow-phase sweep yields a 31.5\% speedup at 2$\times$ scale.

\paragraph{Hot-path optimizations.}
Profiling identified two critical bottlenecks.  First,
\texttt{std::pow(rmag, 5.0)} in the J2 acceleration was 10--50$\times$
slower than explicit multiplication (\texttt{r2*r2*rmag}); replacing it
eliminated the dominant per-object cost.  Second, the three J2 secular
rate functions (\S\ref{sec:methods}) shared common subexpressions
($\cos i$, $\sin i$, $a^{3.5}$) that were being recomputed independently;
factoring them into a precomputed struct removed redundant
transcendental evaluations.  Third, the MOID grid resolution was increased
from 24 to 72~samples after analysis showed that $15^{\circ}$~angular
spacing left $\sim$1{,}780\,km arc gaps at typical LEO altitudes,
rendering the filter effectively inert.

\subsection{Deployment Verification}

The Docker API smoke gate (\texttt{scripts/docker\_api\_smoke.sh}) performs
a fully automated release-verification cycle:

\begin{enumerate}[nosep,leftmargin=*]
  \item Build Docker image from \texttt{Dockerfile}
  \item Boot container on auto-selected port
  \item Probe \texttt{/api/status}, POST telemetry, POST simulate/step,
    GET snapshot
  \item Validate each response against expected schemas with Python
    assertions
  \item Tear down container
\end{enumerate}

\noindent
This end-to-end gate catches regressions that unit tests alone
cannot: image-build failures, port-binding issues, static-asset
serving, and CORS configuration.

%% ═══════════════════════════════════════════════════════════════════════════
\section{Deployment}\label{sec:deploy}

\subsection{Docker Build Architecture}

CASCADE uses a 3-stage multi-stage Dockerfile designed for reproducibility,
build speed, and minimal runtime footprint:

\begin{table}[h]
\centering\small
\begin{tabularx}{\linewidth}{@{}llX@{}}
\toprule
\textbf{Stage} & \textbf{Base} & \textbf{Purpose} \\
\midrule
builder  & \texttt{ubuntu:22.04} & GCC-12, CMake, sccache, CTest gates \\
frontend & \texttt{node:24-slim}  & \texttt{npm ci} + Vite production build \\
runtime  & \texttt{ubuntu:22.04} & Minimal: \texttt{libgomp1} + binary \\
\bottomrule
\end{tabularx}
\caption{Docker build stages.}
\label{tab:docker}
\end{table}

\noindent
Several optimizations accelerate the build pipeline:

\begin{itemize}[nosep,leftmargin=*]
  \item \textbf{sccache}: Mozilla's distributed compiler cache (v0.8.1,
    static musl binary) is used as \texttt{CMAKE\_C\_COMPILER\_LAUNCHER},
    eliminating redundant recompilation across incremental builds.
  \item \textbf{Dependency pre-fetch}: A dedicated Docker layer copies
    only \texttt{CMakeLists.txt} and runs with
    \texttt{PROJECTBONK\_PREFETCH\_ONLY=ON}, downloading simdjson and
    cpp-httplib into a cache-mounted \texttt{\_deps/} directory.  Source
    changes therefore never re-trigger dependency downloads.
  \item \textbf{BuildKit cache mounts}: Four persistent caches (apt,
    \texttt{\_deps}, build directory, sccache) survive across builds.
  \item \textbf{Build-time testing}: \texttt{ctest} runs inside the
    builder stage, meaning the Docker image \emph{cannot be built} if any
    safety gate fails.
\end{itemize}

\noindent
The runtime image runs as a non-root user (\texttt{bonk}), copies only the
compiled binary, ground-station CSV, and the Vite-built frontend static
assets.  CORS is enabled via \texttt{PROJECTBONK\_CORS\_ENABLE=1} for
frontend integration.  The binary serves the frontend directly from
\texttt{cpp-httplib}'s static mount point---no Nginx or reverse proxy
required.

\subsection{Quick Start}

\begin{lstlisting}[language=bash]
# Build and run (entire pipeline, including tests)
docker build -t cascade:local .
docker run --rm -p 8000:8000 cascade:local

# Verify the engine is healthy
curl -s http://localhost:8000/api/status \
  | python3 -m json.tool

# Full automated verification
./scripts/docker_api_smoke.sh
./scripts/run_ready_demo.sh
\end{lstlisting}

\subsection{CI Pipeline}

GitHub Actions runs a build-and-gate workflow on every push to
\texttt{main} and every pull request targeting \texttt{main}:

\begin{enumerate}[nosep,leftmargin=*]
  \item Install toolchain and configure CMake (\texttt{gcc-12}, Release,
    \texttt{BUILD\_TESTING=ON})
  \item Build submission targets and all core gate binaries
  \item Run dedicated shell gate scripts for propagation, broad phase,
    narrow phase, recovery, maneuver ops, and API-contract validation
  \item Run the CTest safety suite with \texttt{--output-on-failure}
  \item Run the Docker API smoke gate, which rebuilds the packaged image
    and verifies backend plus bundled frontend startup
\end{enumerate}

\noindent
Any failure blocks the workflow, ensuring that revisions merged into
\texttt{main} have already passed both source-level gate checks and the
packaged-runtime smoke verification.

%% ═══════════════════════════════════════════════════════════════════════════
\section{Engineering Scope and Future Extensions}

Every engineering project involves deliberate scoping decisions.  The
following items represent conscious trade-offs we made to deliver a
reliable, well-tested system within the hackathon timeline, along with
the extensions we would pursue given additional time.

\begin{itemize}[nosep,leftmargin=*]
  \item \textbf{Perturbation model.}
    The engine models J2 oblateness, which dominates all other
    perturbations in LEO by 2--3 orders of magnitude over the
    problem statement's days-level prediction horizon.  Atmospheric drag
    is applied heuristically for objects with perigee below 350\,km.
    Lunar/solar third-body effects and solar radiation pressure are not
    currently modeled; incorporating them would improve accuracy at long
    horizons but adds complexity disproportionate to the benefit at the
    24-hour CDM timescale.

  \item \textbf{Maneuver optimization.}
    The COLA planner explores a discrete 160-candidate space
    (4~magnitudes $\times$ 4~timings $\times$ 10~directions).  A
    continuous nonlinear programming or sequential convex programming
    formulation would yield lower-$\Delta V$ evasion sequences, but the
    discrete search completes in $<$\,1\,ms, produces deterministic and
    fully auditable results, and---crucially---admits exhaustive
    safety verification of every candidate.  The same determinism that
    enables reproducible CI gates would be harder to maintain with a
    continuous optimizer.

  \item \textbf{Shadow-mode filter activation.}
    The D-criterion, plane/phase, and MOID gates currently operate in
    shadow mode, accumulating calibration evidence.  Activating them would
    reduce narrow-phase compute by an estimated 40--60\% with no
    impact on safety, once the calibration data confirms zero false
    negatives across adversarial edge cases.  The infrastructure for
    activation is fully built---it requires only flipping an environment
    variable per filter.

  \item \textbf{Dense-scene rendering.}
    Frontend rendering at 10{,}000+ debris markers has been verified in
    development; formal 60\,FPS proof under worst-case load is an
    outstanding validation item.  The Canvas~2D rendering path is
    inherently scalable (no DOM nodes per marker), so we expect this
    to be confirmed with a targeted benchmark.
\end{itemize}

%% ═══════════════════════════════════════════════════════════════════════════
\section*{Acknowledgments}

We thank the organizers of the National Space Hackathon 2026 at IIT~Delhi
for designing a technically rigorous and operationally grounded problem
statement that pushed us to engage deeply with the real challenges of
autonomous spacecraft operations.  Space-Track.org provided the real OMM
catalog used for full-scale validation.  The GCC, simdjson, cpp-httplib,
React, Three.js, and Vite open-source communities provided the
foundational tools on which CASCADE was built.

%% ═══════════════════════════════════════════════════════════════════════════
\appendix

\section{Key Physical Constants}\label{app:constants}

\begin{table}[h]
\centering\small
\begin{tabular}{@{}llr@{}}
\toprule
\textbf{Symbol} & \textbf{Description} & \textbf{Value} \\
\midrule
$\mu$       & Earth grav.\ parameter  & 398\,600.4418\,km$^3$/s$^2$ \\
$R_E$       & Earth equatorial radius  & 6\,378.137\,km \\
$J_2$       & Oblateness coefficient   & $1.08263 \times 10^{-3}$ \\
$d_{\text{coll}}$ & Collision threshold & 0.100\,km \\
$I_{\text{sp}}$ & Specific impulse      & 300\,s \\
$g_0$       & Standard gravity         & 9.80665\,m/s$^2$ \\
$m_{\text{dry}}$ & Satellite dry mass   & 500\,kg \\
$m_{\text{fuel}}$ & Initial propellant  & 50\,kg \\
$|\Delta V|_{\max}$ & Thrust limit     & 15\,m/s \\
$\tau_{\text{cd}}$ & Cooldown period    & 600\,s \\
$m_{\text{EOL}}$ & Fuel EOL guard       & 2.5\,kg \\
$\tau_{\text{sig}}$ & Signal latency    & 10\,s \\
$r_{\text{box}}$ & Station-keeping box  & 10\,km \\
\bottomrule
\end{tabular}
\caption{System constants (from \texttt{src/types.hpp}).}
\label{tab:constants}
\end{table}

\section{CDM Severity Tiers}\label{app:cdm}

\begin{table}[h]
\centering\small
\begin{tabular}{@{}llc@{}}
\toprule
\textbf{Tier} & \textbf{Condition} & \textbf{Auto-COLA} \\
\midrule
CRITICAL & $d < 0.100$\,km  & Yes \\
WARNING  & $d < 1.0$\,km    & No (monitor) \\
WATCH    & $d < 5.0$\,km    & No (log)     \\
\bottomrule
\end{tabular}
\caption{Conjunction severity classification.}
\label{tab:cdm}
\end{table}

\section{Verification Commands}\label{app:cmds}

\begin{lstlisting}[language=bash]
# Build (local)
cmake -S . -B build -DCMAKE_BUILD_TYPE=Release
cmake --build build --target ProjectBONK

# Run all safety gates
ctest --test-dir build --output-on-failure

# Individual gates
./scripts/api_contract_gate.sh ./build
./scripts/broad_phase_sanity_gate.sh
./scripts/narrow_phase_false_negative_gate.sh
./scripts/recovery_planner_invariants_gate.sh

# Docker release verification
./scripts/docker_api_smoke.sh

# Frontend build
cd frontend && npm ci && npm run build

# Ready demo
./scripts/run_ready_demo.sh
\end{lstlisting}

\section{NSGA-II Offline Tuner}\label{app:tuner}

Configuring a multi-tier screening pipeline with 69~free parameters requires
more than manual trial-and-error.  We invested in a principled offline
optimization tool: a full NSGA-II (Deb~et~al., 2002) multi-objective
evolutionary optimizer implemented in Julia and bridged to the C++ engine
via \texttt{jluna}.  The optimizer simultaneously minimizes three objectives:

\begin{enumerate}[nosep,leftmargin=*]
  \item \textbf{Safety risk} (number of false negatives; any $>0$ renders
    the solution infeasible via feasibility-first constraint handling)
  \item \textbf{Fuel consumption} (total $\Delta V$ across the
    constellation over the evaluation scenario)
  \item \textbf{Tick latency} (wall-clock time for one simulation step)
\end{enumerate}

\noindent
The 69~parameters span three tiers:

\begin{itemize}[nosep,leftmargin=*]
  \item \textbf{Tier 1 --- Broad phase} (9~params): SMA bin width,
    inclination bin width, shell margin, neighbor search range,
    D-criterion threshold, fail-open eccentricity cutoff.
  \item \textbf{Tier 2 --- Narrow phase} (21~params): plane/phase
    angular thresholds, MOID grid resolution and reject threshold,
    TCA guard margin, micro-refinement band, full-refine budget
    tiers, uncertainty promotion velocity cutoff.
  \item \textbf{Tier 3 --- Maneuver planner} (39~params): COLA
    $\Delta V$ scales, RTN direction weights, CW horizon bounds,
    position/velocity correction blend, recovery budget cap,
    CDM scanner substep and horizon.
\end{itemize}

\noindent
A representative run (20~population $\times$ 5~generations =
120~evaluations) confirmed that all Pareto-optimal candidates maintained
safety\_risk $= 0$, with the best tick latency at 13.3\,ms versus 23\,ms
under na\"ive defaults---a 42\% improvement with zero safety compromise.
The tuner's output became the production defaults that the engine ships
with.  All~69 parameters remain overridable at runtime via
\texttt{PROJECTBONK\_*} environment variables, enabling rapid
experimentation without recompilation.

\section{Problem Statement Alignment}\label{app:ps}

Table~\ref{tab:ps} maps each PS~\S7 evaluation criterion to the
CASCADE subsystem and evidence that addresses it.

\begin{table}[h]
\centering\small
\begin{tabularx}{\linewidth}{@{}lcX@{}}
\toprule
\textbf{Criterion} & \textbf{Wt.} & \textbf{CASCADE Evidence} \\
\midrule
Safety Score    & 25\% & Zero false negatives (8 gates), fail-open
  escalation, conservative screening margin \\
Fuel Efficiency & 20\% & Tsiolkovsky accounting, CW/ZEM recovery,
  per-burn $\Delta V$ tracking \\
Constellation Uptime & 15\% & 10\,km slot box, J2 slot propagation,
  recovery\_slot\_gate ($\leq 0.08$) \\
Algorithmic Speed & 15\% & 13.3\,ms baseline tick, $O(N\log N)$
  broad phase, OpenMP narrow phase \\
UI/UX \& Viz.  & 15\% & 6-page dashboard, shared tactical shell,
  ground track, bullseye, Gantt timeline \\
Code Quality    & 10\% & 8 CTest gates, CI enforcement, Docker smoke,
  env-override system, debug APIs \\
\bottomrule
\end{tabularx}
\caption{PS evaluation criteria alignment.}
\label{tab:ps}
\end{table}

\section{Development Timeline}\label{app:timeline}

CASCADE was developed over approximately 4~weeks across 60+~commits.
The timeline below reflects a deliberate progression from foundational
infrastructure to validated, deployable software.

\begin{table}[h]
\centering\small
\begin{tabularx}{\linewidth}{@{}lX@{}}
\toprule
\textbf{Phase} & \textbf{Milestone} \\
\midrule
Week~1 & Core infrastructure: telemetry ingestion via simdjson, SoA
  state store, J2 propagator with adaptive mode selection, Kepler
  solver, REST API skeleton on port~8000 \\
Week~2 & Screening pipeline: conservative broad phase (shell overlap,
  D-criterion), multi-gate narrow phase (TCA, MOID, RK4 refinement),
  CW/ZEM recovery solver, Tsiolkovsky fuel accounting, cooldown and
  thrust-cap enforcement \\
Week~3 & Optimization and validation: shadow-gate observability, NSGA-II
  tuner (69~params), real-data scenario generator (27K Space-Track
  objects), OpenMP parallelization, environment-variable override
  system, 8~CTest safety gates \\
Week~4 & Integration and polish: React/Canvas/WebGL dashboard (6~pages),
  operational dashboard design, 3-stage Docker build with sccache,
  CI pipeline, PS alignment analysis, technical report \\
\bottomrule
\end{tabularx}
\caption{Development timeline.  Each week built upon the validated outputs
  of the prior phase.}
\label{tab:timeline}
\end{table}

\end{document}
